---
title: "Perspective Taking Summary"
author: "Jeremy Yeaton"
date: "October 17, 2017"
output: html_document
---

```{r load setup, include = FALSE}
load("persp_taking_environment.RData")
library(xtable)
```

3. Run a two-way ANOVA by considering increase factor score as the response
   variable and Daycare and Age group as explanatory variables, with their
   interaction. (a) Use an interaction plot to summarize the main
   findings. (b) Is the interaction significant at the 5% level? (c) If not,
   summarize the effect of each factor (partial eta squared and p-value).

4. Compare results obtained in (3) with those from a two-way ANOVA relying
   on (a) Type II or (b) Type III sum of squares which are available in the
   Anova() command from the car R package (install.packages("car")). Are results
   influenced by the order in which factors enter the model in each case?
   (See http://goo.gl/JUJNGj for details about computing type I, II and III
   SS.)
   
1. Does daycare experience improve overall performance on perspective
   taking?
2. Is this result influenced by the age of the subject?

Types of ANOVAs:
All 3 types will result in the same interaction, but differ on how they handle main effects
Type I is better for measuring main effects in order
Type III is better for interactions
Type II  only if you think there is no effect

## Summary of data
```{r}
table_daycare
table_age
table_cross
```

## Figures
Use box-and-whiskers charts or density plots to show the distribution of individual increase in the composite score for each treatment.
Box and whisker:
```{r}
treatments.plot
```

Density:
```{r}
density.plot
```

##ANOVAs
```{r}
daycare.ezanova
```
