---
title: 'Assignment #1'
author: "Jeremy Yeaton"
date: "November 25, 2017"
output:
  html_document: default
  pdf_document: default
---
```{r, echo=FALSE}
load("assignment01_environment.RData")
library(knitr)
````

# Question 1:

(a) *Give a numerical summary of FEV1 (mean, standard deviation and range)
   for each smoking category (recoded as a categorical variable with
   appropriate levels), and for all subjects (grand mean and overall
   standard deviation). Results should be printed in one or two Tables.* 
   
   The table below shows the descriptive statistics for each of the groups, as well as a summary for the whole sample directly below.
```{r,echo=F,results='asis'}
kable(descript_tbl)
   
kable(overall_tbl)
````

(b) *Use box-and-whiskers charts or density plots to show the distribution of individual values.*

Below is a box-and-whisker plot of all of the points in the sample.

```{r,echo=F}
box_whisker_summary.plot
````

The following box-and-whisker chart displays the distribution of the individual values by group, and provides a more useful insight into the data than the one above.

```{r,echo=FALSE}
box_whisker_bycat.plot
````
  
# Question 2:
*Carry out a one-way ANOVA to test the null hypothesis that FEV1 does not depend on smoking category.*

```{r,echo=F}
by_cat.anova
````

(a) *Formulate your conclusion in plain English, and*

Based on our ANOVA, it seems to be the case that FEV1 *does* depend on smoking category, with a very low p-value and a high F-value, and as such, we can reject the null-hypothesis.

(b) *report the percentage of explained variance.*

The percentage of variance explained by this test is about 35.2%.
   
# Question 3:
(a) *Use post-hoc Tukey HSD tests (R command: TukeyHSD) to compare all
   pairs of means among the four groups of smokers. Summarize point estimates and 95% confidence intervals in a Table or graphical display,and indicate which pairs of means are found to be significantly different.*
   
```{r,echo=F,fig.height=10,fig.width=6}
summary(f.thsd)
   
thsd
   
plot(thsd)
````

   Based on the Tukey HSD test, it would appear that only the difference between the **current smoker** and **non-smoker** groups is significant, since that is the only pair where the adjusted p-value is less than .05.
   
(b) *Compare those results with results from all pairwise comparisons for
   mean FEV1 using the Bonferroni method (R command: pairwise.t.test).*
   
```{r}
bonferroni.ttest
````

We see the same result here, with only the pair **current/non-smoker** returning a p-value less than .05, and as such, is the only pair where we can assume a significant difference.
   
# Question 4:
*Is there any evidence for a linear or quadratic trend for mean FEV1 when
   considering smoking status as ordered factor levels: 1 < 2 < 3 < 4 (use
   the R command factor with the ordered = TRUE option*
   
# Question 5:
(a) *Compare the preceding results with the conclusion that would be
   reached by using a regression approach where one considers smoking status as a numerical variable, as well as its square, i.e., using the R command lm with a formula like FEV1 ~ smoking + I(smoking)^2.*
   
(b) *What could explain the difference, if any?*

#Appendix

```{r}
f_clean
````
